Milk intake and stroke mortality in the Japan Collaborative Cohort Study - a Bayesian survival analysis

### 2020-07-14 paprfolder_docx/strokeJACC.Rmd
updated docx version

---
title: Milk intake and stroke mortality in the Japan Collaborative Cohort Study - a Bayesian survival analysis 
author:
  - name: Chaochen Wang
    affil: 1,*
    orcid: 0000-0001-5533-1497
  - name: Hiroshi Yatsuya
    affil: 2 
    orcid: 0000-0002-6220-9251
  - name: Yingsong Lin
    affil: 1
    orcid: 0000-0003-0214-3649
  - name: Tae Sasakabe 
    affil: 1 
    orcid: 0000-0002-7417-5693
  - name: Sayo Kawai
    affil: 1
    orcid: 0000-0002-5156-4880
  - name: Shogo Kikuchi
    affil: 1
    orcid: 0000-0001-9499-9668
  - name: Hiroyasu Iso
    affil: 3
    orcid: 0000-0002-9241-7289 
  - name: Akiko Tamakoshi
    affil: 4
    orcid: 0000-0002-9761-3879 
affiliation:
  - num: 1
    address: |
      Department of Public Health, Aichi Medical University School of Medicine, Nagakute, Japan
#    email: chaochen@wangcc.me
  - num: 2
    address: |
      Departmet of Public Health, Fujita Health University School of Medicine, Toyoake, Japan
  - num: 3
    address: |
      Public Health, Department of Social Medicine, Osaka University Graduate School of Medicine, Osaka, Japan
  - num: 4
    address: |
      Department of Public Health, Faculty of Medicine, Hokkaido University, Sapporo, Japan
#    email: yatsuya@gmail.com
# firstnote to eighthnote
# firstnote: |
#   Current address: Updated affiliation
# secondnote: |
#   These authors contributed equally to this work.
correspondence: |
  Email.: chaochen@wangcc.me; Tel.: +81-561-62-3311. Department of Public Health, Aichi Medical University School of Medicine, 1-1 Yazakokarimata, Nagakute, Aichi, 480-1195, Japan (C.W.)
journal: nutrients
type: article
status: To be submitted
bibliography: milk.bib
# appendix: appendix.tex
# simplesummary: |
#  A Simple summary goes here.
abstract: |

@@

keywords: |
  milk intake; mortality; stroke; Bayesian survival anlysis; time-to-event data; JACC study 
acknowledgement: |
  The authors would like to express their sincere appreciate to Kunio Aoki and Yoshiyuki Ohno, Professors Emeritus at Nagoya University School of Medicine and former chairpersons of the JACC Study. The whole member of JACC Study Group can be found at https://publichealth.med.hokudai.ac.jp/jacc/index.html. The JACC Study has been supported by Grants-in-Aid for Scientific Research from the Ministry of Education, Culture, Sports, Science and Technology of Japan (MEXT, Monbu Kagaku-sho), Tokyo (grant numbers 61010076, 62010074, 63010074, 1010068, 2151065, 3151064, 4151063, 5151069, 6279102, 11181101, 17015022, 18014011, 20014026, 20390156, 26293138, and 16H06277), and Grants-in-Aid from the Ministry of Health, Labour and Welfare, Health and Labour Sciences Research Grants, Japan [H20-Junkankitou (Seishuu)-Ippan-013, H23-Junkankitou (Seishuu)- Ippan-005, H26-Junkankitou (Seisaku)-Ippan-001, and H29-Junkankitou (Seishuu)-Ippan-003]. 
authorcontributions: |
  ``C.W. and H.Y. conceive and designed the study; C.W. analyzed the data; C.W. wrote the first draft of the paper. A.T. provided the database. All of the authors approved and finalized the manuscript for publication.''
conflictsofinterest: |
  The authors declare no conflict of interest. The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, an in the decision to publish the results.
# sampleavailability: |
#   Samples of the compounds ...... are available from the authors.
abbreviations:
  - short: JACC
    long: Japan Collaborative Cohort 
  - short: FFQ
    long: Food Frequency Questionnaire
  - short: MCMC
    long: Markov Chain Monte Carlo
  - short: JAGS
    long: Just Another Gibbs Samplers
  - short: AFT
    long: accelerated failure time 
  - short: HR
    long: hazard ratio
  - short: AF
    long: acceleration factor
  - short: sd
    long: standard deviation
  - short: CrI
    long: credible interval
header-includes:
   - \usepackage{graphicx}
output: rticles::mdpi_article
---

@@

```{r tab1f, echo=FALSE, eval=FALSE}
library(knitr)
library(kableExtra)

dt <- read.csv("tab/tab1.csv", header = T)
names(dt) <- c("", "Never", "Drinker", "1-2 times/Month", "1-2 times/Week", "3-4 times/Week", "Almost Daily", "P value")

kable(dt, format = "latex", caption = "Sex-specific baseline characteristics according to the frequency of milk intake.", booktabs = T,linesep = "") %>%
kable_styling(latex_options = c("striped","hold_position"), font_size = 8) %>%
  add_header_above(c(" ", "", "", "Milk drinkers" = 4, "")) %>%
  # add_indent(c(4:6, 8:11, 14:17, 19, 20, 22:25, 27:30, 32:35, 37:39, 41:43, 45, 47, 49, 51, 53, 55)) %>%
  pack_rows("Men (n = 39386)", 1, 16) %>%
  pack_rows("Women (n = 545999)", 17, 32) #%>% 
  # footnote(c("Abbreviations: n, number; SD, standard deviation; Alc_Fre, alcohol intake; BMI, Body Mass Index; Slepgrp, sleep duration", "groups; Veg, vegetable intake; Coffe, coffee intake; DM/HT/KID/LIV_hist, diabetes, hypertension, kidney disease,", "liver disease history."))
```

\begin{table}[h]
\caption{\label{tab:tab1f}Sex-specific baseline characteristics according to the frequency of milk intake (JACC study, 1988-2009).}
\centering
\fontsize{8}{10}\selectfont
\resizebox{\textwidth}{!}{%
\begin{tabular}[t]{lccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{4}{c}{\textbf{Milk drinkers}} & \multicolumn{1}{c}{} \\
\cmidrule(l{3pt}r{3pt}){4-7}
 & \textbf{Never} & \textbf{Drinker} & \textbf{1-2 times/} & \textbf{1-2 times/} & \textbf{3-4 times/} & \textbf{Almost} & \\
 &       &         & \textbf{Month}     & \textbf{Week}      & \textbf{Week}         & \textbf{Daily}  & \textbf{\textit{P} value}\\
\midrule
\rowcolor{gray!6}  \addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Men (n = 39386)}}\\
\hspace{1em}number of subjects & 8508 & 30878 & 3522 & 5928 & 5563 & 15865 & \\
\hspace{1em}Age, year (mean (SD)) & 56.8 (9.9) & 56.8 (10.2) & 55.2 (10.1) & 55.4 (10.1) & 55.4 (9.9) & 58.1 (10.1) & <0.001\\
\rowcolor{gray!6}  \hspace{1em}Current smoker, \% & 58.7 & 49.8 & 57.4 & 55.9 & 51.1 & 45.4 & <0.001\\
\hspace{1em}Daily alcohol drinker, \% & 51.9 & 47.8 & 50.9 & 48.4 & 48.6 & 46.5 & <0.001\\
\rowcolor{gray!6}  \hspace{1em}BMI, kg/m$^2$ (mean (SD)) & 22.6 (3.4) & 22.7 (3.4) & 22.8 (2.8) & 22.8 (2.8) & 22.9 (5.4) & 22.6 (2.8) & <0.001\\
\hspace{1em}Exercise (> 1h/week), \% & 19.0 & 27.6 & 26.5 & 25.0 & 25.5 & 29.5 & <0.001\\
\rowcolor{gray!6}  \hspace{1em}Sleep duration, 8-9 hours, \% & 35.6 & 35.9 & 34.6 & 36.2 & 35.1 & 36.3 & <0.001\\
\hspace{1em}Vegetable intake, daily, \% & 21.3 & 25.4 & 20.1 & 20.4 & 20.8 & 30.1 & <0.001\\
\rowcolor{gray!6}  \hspace{1em}Fruit intake, daily, \% & 14.8 & 22.4 & 15.4 & 16.3 & 17.3 & 28.1 & <0.001\\
\hspace{1em}Green tea intake, daily, \% & 76.5 & 79.2 & 79.9 & 78.3 & 77.9 & 79.8 & <0.001\\
\rowcolor{gray!6}  \hspace{1em}Coffee intake, daily, \% & 43.8 & 50.7 & 50.5 & 48.0 & 47.5 & 52.9 & <0.001\\
\hspace{1em}Educated over 18 years old, \% & 25.5 & 34.7 & 33.8 & 33.3 & 31.0 & 36.6 & <0.001\\
\rowcolor{gray!6}  \hspace{1em}History of diabetes, \% & 5.5 & 6.3 & 4.5 & 4.2 & 5.5 & 7.7 & <0.001\\
\hspace{1em}History of hypertension, \% & 18.4 & 17.9 & 17.5 & 17.1 & 16.8 & 18.7 & 0.039\\
\rowcolor{gray!6}  \hspace{1em}History of kidney diseases, \% & 3.0 & 3.4 & 3.8 & 3.0 & 3.0 & 3.5 & <0.001\\
\hspace{1em}History of liver diseases, \% & 5.8 & 6.5 & 6.3 & 6.0 & 5.4 & 7.2 & <0.001\\
\rowcolor{gray!6}  \addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Women (n = 545999)}}\\
\hspace{1em}number of subjects & 10407 & 44592 & 3640 & 7590 & 8108 & 25254 & \\
\hspace{1em}Age, year (mean (SD)) & 58.0 (10.2) & 56.9 (9.9) & 56.5 (10.2) & 55.6 (10.1) & 55.6 (9.9) & 57.9 (9.9) & <0.001\\
\rowcolor{gray!6}  \hspace{1em}Current smoker, \% & 6.9 & 4.2 & 6.1 & 5.5 & 4.3 & 3.5 & <0.001\\
\hspace{1em}Daily alcohol drinker, \% & 4.3 & 4.5 & 5.5 & 4.3 & 4.2 & 4.6 & <0.001\\
\rowcolor{gray!6}  \hspace{1em}BMI, kg/m2 (mean (SD)) & 23.0 (3.4) & 22.9 (3.7) & 23.0 (3.8) & 23.1 (4.4) & 23.1 (3.1) & 22.8 (3.6) & <0.001\\
\hspace{1em}Exercise (> 1h/week), \% & 13.6 & 20.8 & 17.1 & 18.5 & 18.8 & 22.6 & <0.001\\
\rowcolor{gray!6}  \hspace{1em}Sleep duration, 8-9 hours, \% & 27.7 & 25.6 & 25.1 & 25.9 & 25.4 & 25.7 & <0.001\\
\hspace{1em}Vegetable intake, daily, \% & 24.7 & 30.4 & 25.0 & 24.6 & 24.2 & 34.8 & <0.001\\
\rowcolor{gray!6}  \hspace{1em}Fruit intake, daily, \% & 25.0 & 35.7 & 26.6 & 29.2 & 29.2 & 41.1 & <0.001\\
\hspace{1em}Green tea intake, daily, \% & 73.8 & 76.8 & 77.0 & 76.4 & 75.8 & 77.3 & <0.001\\
\rowcolor{gray!6}  \hspace{1em}Coffee intake, daily, \% & 39.6 & 48.2 & 46.2 & 46.4 & 44.4 & 50.2 & <0.001\\
\hspace{1em}Educated over 18 years old, \% & 19.9 & 31.6 & 27.9 & 29.8 & 27.4 & 34.0 & <0.001\\
\rowcolor{gray!6}  \hspace{1em}History of diabetes, \% & 2.6 & 3.7 & 3.2 & 2.7 & 2.7 & 4.4 & <0.001\\
\hspace{1em}History of hypertension, \% & 21.5 & 19.7 & 20.5 & 19.1 & 18.9 & 20.0 & <0.001\\
\rowcolor{gray!6}  \hspace{1em}History of kidney diseases, \% & 3.6 & 4.1 & 3.9 & 3.7 & 3.7 & 4.4 & <0.001\\
\hspace{1em}History of liver diseases, \% & 3.5 & 4.6 & 4.9 & 3.9 & 3.9 & 5.0 & <0.001\\
\bottomrule
\end{tabular}}
\end{table}

@@

As listed in \textbf{Table \ref{tab:tab1f}}, compared with those who chose "never" as their milk intake frequency at baseline, milk drinkers were less likely to be a current smoker or a daily alcohol consumer in both men and women. Furthermore, people consumed milk more than 1-2 times/month were more likely to be a daily consumers of vegetable, fruit  as well as coffee, and more likely to join exercise more than 1 hour/week among both sex.

Detailed results from the Bayesian survival models (crude, age-adjusted and multivariable-adjusted) according to the frequency of milk intake separated by sex are listed in \textbf{Table \ref{tab:tab2}} (men) and \textbf{Table \ref{tab:tab3}} (women). Compared to those who never had milk, both men and women had slower speed and lower hazard of dying from total stroke in crude models. Velocities that milk consumers dying from stroke is slower by a crude acceleration factor (AF) between 0.79 (sd = 0.05; 95% CrI: 0.74, 0.90) and 0.93 (sd = 0.04; 95% CrI: 0.85, 1.02) compared with non-consumers. Chances that the posterior crude HRs were estimated to be lower than 1 for those who had at least 1-2 times/month was higher than 86.5% in men and greater than 94.6% in women. However, lower hazard and delayed time-to-event was observed to remain after age or multivariable adjustment only among daily male milk consumers. Specifically, the mean (sd; 95% CrI) of posterior multivariable-adjusted AF and HR for daily male consumers of milk were 0.88 (sd = 0.05; 95% CrI: 0.81, 0.96) and 0.80 (sd = 0.07; 95% CrI: 0.69, 0.93) with a probability of 99.0% to be smaller than the null value (=1). Daily female milk consumers had posterior AFs and HRs that was distributed with means of 0.97 (sd = 0.09; 95% CrI: 0.88, 1.10) and 0.95 (sd = 0.12; 95% CrI: 0.80, 1.17) which had about 78.0% of chance that their HRs could be smaller than 1.

@@

```{r tab2, echo=FALSE, eval=FALSE, message=FALSE}
library(knitr)
library(kableExtra)
dt <- read.csv("tab/HRAFTweib.csv", header = T)
# dt <- data.frame(`Stroke` = c(rep("Total stroke", 15)), dt)
names(dt) <- c(" ", "Never", "1-2 times/Month", "1-2 times/Week", "3-4 times/Week", "Almost Daily")

kable(dt, format = "latex", caption = "Summary of posterior Acceleration Factors (AF) and Hazard Ratios (HR) of mortality from total stroke, different stroke type according to the frequency of milk intake in men (JACC study, 1988-2009).", booktabs = T, linesep = "", align = "lccccc") %>%
  row_spec(0, bold = T) %>% 
    # collapse_rows(columns = 1, valign = "top") %>% 
    pack_rows("Model 0", 4, 8) %>%
  pack_rows("Model 1", 9, 13) %>%
  pack_rows("Model 2", 14, 18) %>%
  row_spec(18,  hline_after = TRUE) %>%
  pack_rows("Model 0", 20, 24) %>%
  pack_rows("Model 1", 25, 29) %>%
  pack_rows("Model 2", 30, 34) %>%
  row_spec(34,  hline_after = TRUE) %>%
  pack_rows("Model 0", 36, 40) %>%
  pack_rows("Model 1", 41, 45) %>%
  pack_rows("Model 2", 46, 50) %>%
kable_styling(latex_options = c("striped","hold_position"), font_size = 7) %>%
  # add_header_above(c(" ", "Estimated HR in Men" = 4, "Estimated HR in Women" = 4)) %>% 
  footnote(c("Abbreviations: SD, standard deviation; CrI, credible interval; MCSE, Monte Carlo Standard Error;", " Pr(HR < 1) indicates the prabability for posterior HR to be smaller than 1.", "Model 0 = Crude model; Model 1 = age-adjusted model; Model 2 = multivariable adjusted model.", "Covariates included in Model 2: age, smoking habit, alcohol intake, body mass index, history of",  "hypertension, diabetes, kidney/liver diseases, exercise, sleep duration, coffee intake, education level.")) 
```

\begin{table}[!h]

\caption{\label{tab:tab2}Summary of posterior Acceleration Factors (AF) and Hazard Ratios (HR) of mortality from total stroke, stroke types according to the frequency of milk intake in men (JACC study, 1988-2009).}
\centering
\fontsize{6}{8}\selectfont
\resizebox*{!}{\dimexpr\textheight-\lineskip\relax}{%
\begin{tabular}[t]{lccccc}
\toprule
\textbf{ } & \textbf{Never} & \textbf{1-2 times/Month} & \textbf{1-2 times/Week} & \textbf{3-4 times/Week} & \textbf{Almost Daily}\\
\midrule
\rowcolor{gray!6}  Person-year & 135704 & 56551 & 97098 & 92153 & 252364\\
N & 8508 & 3522 & 5928 & 5563 & 15865\\
\rowcolor{gray!6}  Total Stroke & 326 & 122 & 181 & 177 & 546\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 0}}\\
\hspace{1em}Mean AF (SD) & 1 & 0.93 (0.07) & 0.83 (0.05) & 0.85 (0.05) & 0.93 (0.04)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.81, 1.06) & (0.73, 0.94) & (0.74, 0.96) & (0.85, 1.02)\\
\hspace{1em}Mean HR (SD) & 1 & 0.89 (0.09) & 0.77 (0.07) & 0.79 (0.07) & 0.90 (0.06)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.73, 1.08) & (0.63, 0.91) & (0.66, 0.94) & (0.79, 1.03)\\
\hspace{1em}Pr(HR < 1) & - & 86.5\% & 99.9\% & 99.7\% & 93.5\%\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 1}}\\
\rowcolor{gray!6}  \hspace{1em}Mean AF (SD) & 1 & 0.99 (0.06) & 0.90 (0.05) & 0.91 (0.05) & 0.85 (0.04)\\
\hspace{1em}95\% CrI & - & (0.87, 1.11) & (0.81, 1.00) & (0.82, 1.01) & (0.78, 0.92)\\
\rowcolor{gray!6}  \hspace{1em}Mean HR (SD) & 1 & 0.98 (0.11) & 0.84 (0.08) & 0.86 (0.08) & 0.76 (0.05)\\
\hspace{1em}95\% CrI & - & (0.79, 1.19) & (0.70, 1.00) & (0.71, 1.02) & (0.66, 0.87)\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 58.7\% & 97.3\% & 96.1\% & 100.0\%\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 2}}\\
\hspace{1em}Mean AF (SD) & 1 & 1.00 (0.07) & 0.92 (0.06) & 0.94 (0.06) & 0.88 (0.05)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.88, 1.14) & (0.82, 1.03) & (0.84, 1.05) & (0.81, 0.96)\\
\hspace{1em}Mean HR (SD) & 1 & 1.01 (0.12) & 0.87 (0.09) & 0.90 (0.09) & 0.80 (0.07)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.81, 1.24) & (0.72, 1.05) & (0.74, 1.08) & (0.69, 0.93)\\
\hspace{1em}Pr(HR < 1) & - & 50.6\% & 93.7\% & 89.6\% & 99.0\%\\
\hline
\rowcolor{gray!6}  Hemorrhagic stroke & 100 & 42 & 58 & 56 & 176\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 0}}\\
\hspace{1em}Mean AF (SD) & 1 & 1.03 (0.17) & 0.85 (0.12) & 0.87 (0.13) & 0.98 (0.11)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.74, 1.38) & (0.63, 1.12) & (0.65, 1.14) & (0.78, 1.22)\\
\hspace{1em}Mean HR (SD) & 1 & 1.03 (0.19) & 0.82 (0.14) & 0.84 (0.15) & 0.97 (0.13)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.70, 1.46) & (0.56, 1.14) & (0.60, 1.17) & (0.75, 1.26)\\
\hspace{1em}Pr(HR < 1) & - & 47.2\% & 88.4\% & 86.3\% & 63.1\%\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 1}}\\
\rowcolor{gray!6}  \hspace{1em}Mean AF (SD) & 1 & 1.08 (0.17) & 0.91 (0.13) & 0.92 (0.13) & 0.90 (0.10)\\
\hspace{1em}95\% CrI & - & (0.80, 1.45) & (0.70, 1.20) & (0.71, 1.19) & (0.74, 1.11)\\
\rowcolor{gray!6}  \hspace{1em}Mean HR (SD) & 1 & 1.11 (0.21) & 0.88 (0.16) & 0.90 (0.16) & 0.88 (0.12)\\
\hspace{1em}95\% CrI & - & (0.75, 1.58) & (0.63, 1.25) & (0.63, 1.24) & (0.67, 1.14)\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 31.6\% & 79.7\% & 76.6\% & 87.6\%\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 2}}\\
\hspace{1em}Mean AF (SD) & 1 & 1.11 (0.18) & 0.93 (0.15) & 0.96 (0.16) & 0.96 (0.13)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.79, 1.58) & (0.70, 1.25) & (0.71, 1.34) & (0.76, 1.25)\\
\hspace{1em}Mean HR (SD) & 1 & 1.14 (0.22) & 0.92 (0.17) & 0.95 (0.18) & 0.95 (0.14)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.75, 1.61) & (0.63, 1.29) & (0.65, 1.37) & (0.71, 1.27)\\
\hspace{1em}Pr(HR < 1) & - & 28.8\% & 72.4\% & 64.4\% & 69.3\%\\
\hline
\rowcolor{gray!6}  Cerebral infarction & 151 & 41 & 64 & 66 & 198\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 0}}\\
\hspace{1em}Mean AF (SD) & 1 & 0.76 (0.09) & 0.71 (0.07) & 0.74 (0.08) & 0.79 (0.06)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.59, 0.94) & (0.58, 0.86) & (0.61, 0.89) & (0.68, 0.93)\\
\hspace{1em}Mean HR (SD) & 1 & 0.65 (0.12) & 0.59 (0.09) & 0.64 (0.09) & 0.71 (0.09)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.46, 0.92) & (0.43, 0.79) & (0.47, 0.85) & (0.56, 0.89)\\
\hspace{1em}Pr(HR < 1) & - & 99.1\% & 99.9\% & 99.7\% & 99.5\%\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 1}}\\
\rowcolor{gray!6}  \hspace{1em}Mean AF (SD) & 1 & 0.83 (0.08) & 0.79 (0.07) & 0.82 (0.07) & 0.74 (0.05)\\
\hspace{1em}95\% CrI & - & (0.68, 1.01) & (0.67, 0.93) & (0.69, 0.96) & (0.66, 0.84)\\
\rowcolor{gray!6}  \hspace{1em}Mean HR (SD) & 1 & 0.73 (0.13) & 0.65 (0.10) & 0.70 (0.11) & 0.58 (0.07)\\
\hspace{1em}95\% CrI & - & (0.49, 1.02) & (0.48, 0.88) & (0.51, 0.94) & (0.46, 0.72)\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 96.9\% & 99.8\% & 98.9\% & 100.0\%\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 2}}\\
\hspace{1em}Mean AF (SD) & 1 & 0.84 (0.09) & 0.80 (0.08) & 0.83 (0.08) & 0.75 (0.06)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.67, 1.02) & (0.67, 0.95) & (0.69, 0.99) & (0.66, 0.85)\\
\hspace{1em}Mean HR (SD) & 1 & 0.73 (0.14) & 0.67 (0.11) & 0.72 (0.12) & 0.61 (0.08)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.50, 1.04) & (0.48, 0.91) & (0.52, 0.99) & (0.48, 0.79)\\
\hspace{1em}Pr(HR < 1) & - & 96.1\% & 99.1\% & 97.5\% & 99.8\%\\
\bottomrule
\multicolumn{6}{l}{\textit{Note: }}\\
\multicolumn{6}{l}{Abbreviations: SD, standard deviation; CrI, credible interval; MCSE, Monte Carlo Standard Error;}\\
\multicolumn{6}{l}{ Pr(HR < 1) indicates the prabability for posterior HR to be smaller than 1.}\\
\multicolumn{6}{l}{Model 0 = Crude model; Model 1 = age-adjusted model; Model 2 = multivariable adjusted model.}\\
\multicolumn{6}{l}{Covariates included in Model 2: age, smoking habit, alcohol intake, body mass index, history of}\\
\multicolumn{6}{l}{hypertension, diabetes, kidney/liver diseases, exercise, sleep duration, coffee intake, education level.}\\
\end{tabular}}
\end{table}

```{r tab3, echo=FALSE, eval=FALSE, message=FALSE}
library(knitr)
library(kableExtra)
dt <- read.csv("tab/HRAFTweibF.csv", header = T)
# dt <- data.frame(`Stroke` = c(rep("Total stroke", 15)), dt)
names(dt) <- c(" ", "Never", "1-2 times/Month", "1-2 times/Week", "3-4 times/Week", "Almost Daily")

kable(dt, format = "latex", caption = "Summary of posterior Acceleration Factors (AF) and Hazard Ratios (HR) of mortality from total stroke, stroke type according to the frequency of milk intake in women (JACC study, 1988-2009).", booktabs = T, linesep = "", align = "lccccc") %>%
    # collapse_rows(columns = 1, valign = "top") %>% 
  row_spec(0, bold = T) %>% 
    # collapse_rows(columns = 1, valign = "top") %>% 
    pack_rows("Model 0", 4, 8) %>%
  pack_rows("Model 1", 9, 13) %>%
  pack_rows("Model 2", 14, 18) %>%
  row_spec(18,  hline_after = TRUE) %>%
  pack_rows("Model 0", 20, 24) %>%
  pack_rows("Model 1", 25, 29) %>%
  pack_rows("Model 2", 30, 34) %>%
  row_spec(34,  hline_after = TRUE) %>%
  pack_rows("Model 0", 36, 40) %>%
  pack_rows("Model 1", 41, 45) %>%
  pack_rows("Model 2", 46, 50) %>%
kable_styling(latex_options = c("striped","hold_position"), font_size = 7) %>%
  # add_header_above(c(" ", "Estimated HR in Men" = 4, "Estimated HR in Women" = 4)) %>% 
  footnote(c("Abbreviations: SD, standard deviation; CrI, credible interval; MCSE, Monte Carlo Standard Error;", " Pr(HR < 1) indicates the prabability for posterior HR to be smaller than 1.", "Model 0 = Crude model; Model 1 = age-adjusted model; Model 2 = multivariable adjusted model.", "Covariates included in Model 2: age, smoking habit, alcohol intake, body mass index, history of",  "hypertension, diabetes, kidney/liver diseases, exercise, sleep duration, coffee intake, education level.")) 
```

\begin{table}[!h]

\caption{\label{tab:tab3}Summary of posterior Acceleration Factors (AF) and Hazard Ratios (HR) of mortality from total stroke, stroke type according to the frequency of milk intake in women (JACC study, 1988-2009).}
\centering
\fontsize{7}{9}\selectfont
\resizebox*{!}{\dimexpr\textheight-\lineskip\relax}{%
\begin{tabular}[t]{lccccc}
\toprule
\textbf{ } & \textbf{Never} & \textbf{1-2 times/Month} & \textbf{1-2 times/Week} & \textbf{3-4 times/Week} & \textbf{Almost Daily}\\
\midrule
\rowcolor{gray!6}  Person-year & 173222 & 59904 & 129233 & 139919 & 418925\\
N & 10407 & 3640 & 7590 & 8108 & 25254\\
\rowcolor{gray!6}  Total Stroke & 300 & 84 & 182 & 172 & 585\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 0}}\\
\hspace{1em}Mean AF (SD) & 1 & 0.88 (0.07) & 0.87 (0.05) & 0.79 (0.05) & 0.88 (0.04)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.75, 1.03) & (0.78, 0.98) & (0.71, 0.90) & (0.80, 0.96)\\
\hspace{1em}Mean HR (SD) & 1 & 0.83 (0.10) & 0.81 (0.08) & 0.70 (0.07) & 0.81 (0.07)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.64, 1.05) & (0.68, 0.97) & (0.58, 0.85) & (0.71, 0.93)\\
\hspace{1em}Pr(HR < 1) & - & 94.6\% & 98.7\% & 99.9\% & 99.6\%\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 1}}\\
\rowcolor{gray!6}  \hspace{1em}Mean AF (SD) & 1 & 0.99 (0.09) & 1.11 (0.08) & 1.02 (0.08) & 0.95 (0.06)\\
\hspace{1em}95\% CrI & - & (0.85, 1.17) & (0.97, 1.26) & (0.89, 1.16) & (0.86, 1.06)\\
\rowcolor{gray!6}  \hspace{1em}Mean HR (SD) & 1 & 1.00 (0.14) & 1.18 (0.14) & 1.03 (0.12) & 0.92 (0.09)\\
\hspace{1em}95\% CrI & - & (0.76, 1.31) & (0.95, 1.47) & (0.82, 1.28) & (0.78, 1.09)\\
\rowcolor{gray!6}  \hspace{1em}Pr(HR < 1) & - & 52.3\% & 6.3\% & 42.0\% & 86.8\%\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 2}}\\
\hspace{1em}Mean AF (SD) & 1 & 1.01 (0.12) & 1.11 (0.14) & 1.02 (0.12) & 0.97 (0.09)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.85, 1.20) & (0.97, 1.30) & (0.89, 1.19) & (0.88, 1.10)\\
\hspace{1em}Mean HR (SD) & 1 & 1.01 (0.17) & 1.19 (0.15) & 1.03 (0.15) & 0.95 (0.12)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.75, 1.36) & (0.96, 1.52) & (0.81, 1.31) & (0.80, 1.17)\\
\hspace{1em}Pr(HR < 1) & - & 52.8\% & 6.4\% & 44.4\% & 78.0\%\\
\hline
\rowcolor{gray!6}  Hemorrhagic stroke & 108 & 27 & 78 & 76 & 231\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 0}}\\
\hspace{1em}Mean AF (SD) & 1 & 0.78 (0.13) & 0.98 (0.12) & 0.90 (0.11) & 0.92 (0.09)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.55, 1.06) & (0.76, 1.25) & (0.70, 1.13) & (0.76, 1.12)\\
\hspace{1em}Mean HR (SD) & 1 & 0.73 (0.16) & 0.98 (0.15) & 0.87 (0.14) & 0.89 (0.11)\\
\rowcolor{gray!6}  \hspace{1em}95\% CrI & - & (0.47, 1.08) & (0.71, 1.31) & (0.64, 1.16) & (0.71, 1.15)\\
\hspace{1em}Pr(HR < 1) & - & 94.7\% & 58.1\% & 83.1\% & 83.0\%\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Model 1}}\\
\rowcolor{gray!6}  \hspace{1em}Mean AF (SD) & 1 & 0.88 (0.13) & 1.12 (0.13) & 1.04 (0.13) & 0.95 (0.09)\\
\hspace{1em}95\% CrI & - & (0.63, 1.17) & (0.90, 1.41) & (0.82, 1.32) & (0.80, 1.14)\\
\rowcolor{gray!6}  \hspace{1em}Mean HR (SD) & 1 & 0.84 (0.18) & 1.17 (0.18) & 1.06 (0.17) & 0.93 (0.12)\\

>>>>

---
title: "Milk intake and stroke mortality in the Japan Collaborative Cohort Study - a Bayesian survival analysis"
output: word_document
bibliography: ../Paperfolder/milk.bib 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Abstract:

@@

keywords: 
  milk intake; mortality; stroke; Bayesian survival anlysis; time-to-event data; JACC study

@@

@@

As listed in **Table 1**, compared with those who chose "never" as their milk intake frequency at baseline, milk drinkers were less likely to be a current smoker or a daily alcohol consumer in both men and women. Furthermore, people consumed milk more than 1-2 times/month were more likely to be a daily consumers of vegetable, fruit  as well as coffee, and more likely to join exercise more than 1 hour/week among both sex.

Detailed results from the Bayesian survival models (crude, age-adjusted and multivariable-adjusted) according to the frequency of milk intake separated by sex are listed in **Table 2** (men) and **Table 3** (women). Compared to those who never had milk, both men and women had slower speed and lower hazard of dying from total stroke in crude models. Velocities that milk consumers dying from stroke is slower by a crude acceleration factor (AF) between 0.79 (sd = 0.05; 95% CrI: 0.74, 0.90) and 0.93 (sd = 0.04; 95% CrI: 0.85, 1.02) compared with non-consumers. Chances that the posterior crude HRs were estimated to be lower than 1 for those who had at least 1-2 times/month was higher than 86.5% in men and greater than 94.6% in women. However, lower hazard and delayed time-to-event was observed to remain after age or multivariable adjustment only among daily male milk consumers. Specifically, the mean (sd; 95% CrI) of posterior multivariable-adjusted AF and HR for daily male consumers of milk were 0.88 (sd = 0.05; 95% CrI: 0.81, 0.96) and 0.80 (sd = 0.07; 95% CrI: 0.69, 0.93) with a probability of 99.0% to be smaller than the null value (=1). Daily female milk consumers had posterior AFs and HRs that was distributed with means of 0.97 (sd = 0.09; 95% CrI: 0.88, 1.10) and 0.95 (sd = 0.12; 95% CrI: 0.80, 1.17) which had about 78.0% of chance that their HRs could be smaller than 1.

@@
